---
title: "Online Shoppers Intention classification using Boosting"
author: "Nicola Disabato"
date: "2022-08-01"
output:
  github_document:
    html_preview: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing libraries

```{r}
library(magrittr) 
library(dplyr)
library(caret)
library(fastAdaboost)
library(Matrix)
library(ROCR)
library(pROC)
library(xgboost)
library(gbm)
```

## Importing the dataset

```{r}
# Load the dataset and explore
intentions <- read.csv("online_shoppers_intention.csv", header = TRUE) 
str(intentions)
```

```{r}
table(intentions$Revenue)
```

As you can see, the dataset is made up of 12330 instances and 18 features. Of these observations, only 1908 are users who have finalized a purchase.

All details can be found through the following link: <https://archive.ics.uci.edu/ml/datasets/Online+Shoppers+Purchasing+Intention+Dataset#>

## Data Preparation

```{r}
intentions <- intentions %>% 
  mutate(OperatingSystems = as.factor(OperatingSystems),
         Browser = as.factor(Browser),
         Region = as.factor(Region),
         TrafficType = as.factor(TrafficType),
         VisitorType = as.factor(VisitorType),
         Month = as.factor(Month)
         )

intentions <- intentions %>% 
  mutate(
         Weekend = as.numeric(Weekend),
         Revenue = as.numeric(Revenue)
         )

str(intentions)
```

We use one-hot encoding for categorical variables.

```{r}
dmy <- dummyVars(" ~ .", data = intentions)
intentions <- data.frame(predict(dmy, newdata = intentions))

dim(intentions)

```

After the preprocessing phase, we obtain a dataset with 75 features.

## Train and Test partition

```{r}
set.seed(100)
inTrain <- createDataPartition(y = intentions$Revenue, p = .75, list = FALSE)
train <- intentions[ inTrain,] 
test <- intentions[-inTrain,]

X_train <- sparse.model.matrix(Revenue ~ .-1, data = train)
y_train <- train[,"Revenue"]  
X_test <- sparse.model.matrix(Revenue~.-1, data = test)
y_test <- test[,"Revenue"]
```

## Let's explore AdaBoost model

```{r}
model_adaboost <- adaboost(Revenue ~ ., data=train, nIter=10)
model_adaboost
```

After training the model on the train dataset, you can use the predict () method to predict the output of the Revenue class in the test dataset. To analyze the performance of the model, it was decided to print the confusion matrix in addition to the precision, recall and f1-score metrics, in addition to the accuracy metric.
