---
title: "Online Shoppers Intention classification using Boosting"
author: "Nicola Disabato"
date: "2022-08-01"
output:
  github_document:
    html_preview: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing libraries

```{r}
library(magrittr) 
library(dplyr)
library(caret)
library(fastAdaboost)
library(Matrix)
library(ROCR)
library(pROC)
library(xgboost)
library(gbm)
```

## Importing the dataset

```{r}
# Load the dataset and explore
intentions <- read.csv("online_shoppers_intention.csv", header = TRUE) 
str(intentions)
```

```{r}
table(intentions$Revenue)
```

As you can see, the dataset is made up of 12330 instances and 18 features. Of these observations, only 1908 are users who have finalized a purchase.

All details can be found through the following link: <https://archive.ics.uci.edu/ml/datasets/Online+Shoppers+Purchasing+Intention+Dataset#>

## Data Preparation

```{r}
intentions <- intentions %>% 
  mutate(OperatingSystems = as.factor(OperatingSystems),
         Browser = as.factor(Browser),
         Region = as.factor(Region),
         TrafficType = as.factor(TrafficType),
         VisitorType = as.factor(VisitorType),
         Month = as.factor(Month)
         )

intentions <- intentions %>% 
  mutate(
         Weekend = as.numeric(Weekend),
         Revenue = as.numeric(Revenue)
         )

str(intentions)
```

We use one-hot encoding for categorical variables.

```{r}
dmy <- dummyVars(" ~ .", data = intentions)
intentions <- data.frame(predict(dmy, newdata = intentions))

dim(intentions)

```

After the preprocessing phase, we obtain a dataset with 75 features.

## Train and Test partition
